---
title: "Individual 2"
author: "Meredith Tucker (mltucker2001)"
date: "4/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(mapdata)
```

```{r}
acc <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv", stringsAsFactors = FALSE)

pers <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv", stringsAsFactors = FALSE)

head(acc)
```



### Part 1

#1.)are there some days of the week where more accidents happen than on others (use variable DAY_WEEK)?
```{r}
DAY_WEEK <- table(acc$DAY_WEEK)
DAY_WEEK

```


#2.)what time of the day do accidents happen (use variable HOUR)?

```{r}
timeacc <- acc %>% 
  group_by(HOUR) %>% 
  summarize(n=n()) %>%
  arrange(desc(n))
timeacc

```

#What is the number of accidents with at least one drunk driver (use variable DRUNK_DR)?

```{r}
acc %>% 
  group_by(DRUNK_DR) %>% 
  summarize(n=n())
8513+249+7

```


###PART 2

# In the person table, identify drivers( PER_TYP==1, se FARS manual), and subset them
```{r}
Driver <- pers %>% 
  filter(PER_TYP == 1)

Driver
```

# 5.) Join accident and person table
```{r}
accper <- inner_join(acc, Driver)
accper
```

# 6.) Tally the number of accidents by day of the week, hour of the day, and gender. Visualize results.
```{r}
tal1 <- accper %>% 
  group_by(DAY_WEEK) %>% 
  summarize(Weekdaycount=n())

tal2 <- accper %>% 
  group_by(HOUR) %>% 
  summarize(Hourcount=n()) %>% 
  mutate(HOUR = factor(HOUR))

tal3 <- accper %>% 
  group_by(SEX) %>% 
  summarize(sexcount=n()) %>% 
  mutate(SEX = factor(SEX))


tal1
tal2
tal3

ggplot(tal1, aes(x=DAY_WEEK, weight=Weekdaycount)) + geom_bar()

ggplot(tal2, aes(x=HOUR, weight=Hourcount)) + geom_bar()

ggplot(tal3, aes(x=SEX, weight=sexcount)) + geom_bar()

```



### Part 3

# Now plot a choropleth map of accidents on a county level

```{r}
GLCUS <- readxl::read_xlsx('C:/Users/mltuc/Documents/DS 202/ds202_lab5/glc us.xlsx')
head(GLCUS)
```




```{r}
#joining acc and GLCUS data sets
#by state and state code


GLCUS2 <- GLCUS %>% 
  mutate(`State Code` = as.integer(`State Code`), 
         `County Code` = as.integer(`County Code`),
         `City Code` = as.integer(`City Code`)) %>% 
  inner_join(accper, c(`County Code` = 'COUNTY', `State Code` = 'STATE', `City Code` = 'CITY'))

head(GLCUS2)
  

states <- map_data('state')
counties <- map_data('county')

ggplot(data=states, aes(x=long, y=lat)) + 
  geom_polygon(aes(group=group)) + 
  geom_point(data = GLCUS2, aes(x=LONGITUD, y=LATITUDE), color='lightgreen', alpha=2, size=.02) +
  xlim(c(-130, -60)) + ylim(c(20,50))


ggplot(data=counties, aes(x=long, y=lat)) + 
  geom_polygon(data=counties, aes(group=group)) + 
  geom_point(data = GLCUS2, aes(x=LONGITUD, y=LATITUDE), color='lightgreen', alpha=2, size=.02) +
  xlim(c(-130, -60)) + ylim(c(20,50))

```




# Compare Winter verse Summer
```{r}
#Winter
ggplot(data=states, aes(x=long, y=lat)) + 
  geom_polygon(aes(group=group)) + 
  geom_point(data = GLCUS2 %>% filter(MONTH %in% c(12,1,2)), aes(x=LONGITUD, y=LATITUDE), color='lightgreen', alpha=2, size=.02) +
  xlim(c(-130, -60)) + ylim(c(20,50))

#Summer
ggplot(data=states, aes(x=long, y=lat)) + 
  geom_polygon(aes(group=group)) + 
  geom_point(data = GLCUS2 %>% filter(MONTH %in% c(6,7,8)), aes(x=LONGITUD, y=LATITUDE), color='lightgreen', alpha=2, size=.02) +
  xlim(c(-130, -60)) + ylim(c(20,50))
```


