---
title: "Lab 5 team"
author: "Meredith Tucker, (mltucker2001) and Mike Zaino, (mzaino16)"
date: "4/13/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(mapdata)
```


```{r}
acc <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv", stringsAsFactors = FALSE)

pers <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv", stringsAsFactors = FALSE)

head(acc)
```

### Part 1

#1.)are there some days of the week where more accidents happen than on others (use variable DAY_WEEK)?
```{r}
DAY_WEEK <- table(acc$DAY_WEEK)
DAY_WEEK

```

# The days of the week that have the most accidents compared to the other days of the week are Friday, Saturday, and Sunday. Saturdays have the most overall accidents. 


#2.)what time of the day do accidents happen (use variable HOUR)?

```{r}
timeacc <- acc %>% 
  group_by(HOUR) %>% 
  summarize(n=n()) %>%
  arrange(desc(n))
timeacc
```

# The most accidents occur at hour 18 miltary time which is 6pm. This is mostly likely due to end of the work day rush hour.

#What is the number of accidents with at least one drunk driver (use variable DRUNK_DR)?

```{r}
acc %>% 
  group_by(DRUNK_DR) %>% 
  summarize(n=n())
acc
8513+249+7
```

# There are 8769 accidents that involve at least one drunk driver. 


###PART 2

# In the person table, identify drivers( PER_TYP==1, se FARS manual), and subset them

```{r}
Driver <- pers %>% 
  filter(PER_TYP == 1)
```



# 5.) Join accident and person table
```{r}
accper <- inner_join(acc, Driver)
head(accper)
```


# 6.) Tally the number of accidents by day of the week, hour of the day, and gender. Visualize results.

```{r}
tal1 <- accper %>% 
  group_by(DAY_WEEK) %>% 
  summarize(Weekdaycount=n())

tal2 <- accper %>% 
  group_by(HOUR) %>% 
  summarize(Hourcount=n()) %>% 
  mutate(HOUR = factor(HOUR))

tal3 <- accper %>% 
  group_by(SEX) %>% 
  summarize(sexcount=n()) %>% 
  mutate(SEX = factor(SEX))


tal1
tal2
tal3

ggplot(tal1, aes(x=DAY_WEEK, weight=Weekdaycount, fill=DAY_WEEK)) + geom_bar() + ggtitle('Day of Week Count')

ggplot(tal2, aes(x=HOUR, weight=Hourcount)) + geom_bar() + ggtitle('Hour Count')

ggplot(tal3, aes(x=SEX, weight=sexcount)) + geom_bar() + ggtitle('Count by Sex')
```




